{\bf Kernel Methods and SVMs} \\ {\large A convex optimization point of view}

### 2016-06-07 coproject/report_tallec_lederrey.tex
UPD: Finished writing experience + conclusion. Just need to review the whole report

\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{listings}

@@

  In this document, we aim at presenting the general framework of
  Kernel methods, to expose how they relate to convex optimization,
  and how their optimization can be undertaken in practice. Besides,
  we intend to give a more precise treatment of the support vectors
  machines special case.

  Most theoretical results used here are presented, explained and
  proved more thoroughly in the ``Master vision et apprentissage''
  of the ``Ecole normale superieure de Cachan'' course on kernels.
\end{abstract}

\section{Kernel methods}
\subsection{Generalities}
\label{sub:gen}

@@

\section{Support vector machines: a special case}
Support vector machines are kernel methods used in the field of binary
classification, and that can be extended (quite painfully, unfortunately) to
$n$-ary classification.  Among kernel methods, they are probably the best
known.  This might be explained by the fact that they enforce sparsity in the
best fit representation. Formally, they ensure that in eq~\ref{eq:bestfit}, a
varying number of $\alpha$'s will be exactly zero.  Sparsity often brings along
a lot of nice properties. For SVMs, most notably, they ensure a better
regularization (intuitively, a fit with less term is less complex than a fit
with many terms, and the number of term is directly related to the number of
non-zero $\alpha$'s), and they induce smaller computational cost. Indeed, the
best fit represented in eq~\ref{eq:bestfit} has as many terms as there are
datapoints in the train dataset. This means that the computational cost at test
time will depend directly on the number of datapoints in the training set. This
might prove computationally costly. SVMs partly solve this problem by ensuring
that some (most) $\alpha$'s are zero, and thus that the best fit will only
depend on a few training datapoints. 
\subsection{Generalities}
Given an input set $\X$, a p.s.d kernel on $\X$, $K$ and a training dataset,
$((x_1, y_1), \dots, (x_n, y_n)) \in (\X \times \pmset)^n$, SVMs solve the
kernel optimization problem eq~\ref{eq:minkernel} with the particular choice of
loss $L(x, y) = \max(0, 1 - x^\top y)$, which is known as the Hinge Loss.

@@

The theory works well, but a use case is always needed. We decided to use the SVM to learn our computer how to recognize handwritten digits. We used the MNIST database. Some examples are given in Figure \ref{fig:train-set}.

@@

We trained the algorithm with different kernel functions:
\begin{itemize}
\item Linear kernel: $K(x_i,x_j) = \langle x_i, x_j \rangle$
\item Polynomial kernel: $K(x_i,x_j) = (\gamma \langle x_i, x_j \rangle + r)^p$, $\gamma$ is a parameter that needs to be found and we decided to keep $r=0$ and we used $p=4$.
\item RBF kernel:  $K(x_i,x_j) = \exp(-\gamma|x_i-x_j|^2)$, $\gamma$ is a parameter that needs to be found.
\end{itemize}

For each of these kernels, we trained on train-set composed of 60000 handwritten digits with their labels. Then, we test our trained models against a test-set composed of 10000 images. We can then compared the predicted labels against the true labels. 
\\\\
On of the important result is the percentage of correct prediction. It is given for each of the kernel in Table \ref{tab:results}. As we can see, the best kernel is the RBF one. It has already been shown in the literature that this kernel is the best one for SVM.
\begin{table}[H]
\centering
\begin{tabular}{|l|c|}
\hline
Linear Kernel & 94.62\% \\\hline
Polynomial Kernel (degree 4) & 97.31\% \\\hline
RBF Kernel & 98.56 \% \\ \hline
\end{tabular}
\caption{\label{tab:results} Percentage of correct prediction after training.}
\end{table}

{\color{red} Ici, je veux mettre une figure avec toutes les predictions dur RBF. Et finalement montrer quelques fausses predictions. Juste apres, on pourra faire un conclusion.}

>>>>

\usepackage{tikz}
\usepackage{footnote}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{listings}
\usepackage{array,hhline}
\usepackage{pgf}
\usepackage{multirow,graphicx,array,xcolor,colortbl}

\renewcommand*{\arraystretch}{1.5}

\newcommand{\cca}[2]{
	\cellcolor{black!#1}
	\ifnum #1>50
		\color{white}
	\fi
	{#2}
	}

@@

  In this document, we aim at presenting the general framework of
  Kernel methods, to expose how they relate to convex optimization,
  and how their optimization can be undertaken in practice. Besides,
  we intend to give a more precise treatment of the support vectors
  machines special case.
\end{abstract}

\section{Kernel methods\footnotemark}
\subsection{Generalities}
\label{sub:gen}

\footnotetext{The different elements presented in this section comes from the course on Machine Learning \cite{course} as well as the paper explaining how convex optimization is used with Kernel methods\cite{paper}.}

@@

\section{Support vector machines: a special case\footnotemark}
\footnotetext{The different elements presented in this section comes from the course on Machine Learning \cite{course} as well as the article on Wikip\'edia on SVM\cite{svm}.}
Support vector machines are kernel methods used in the field of binary
classification, and that can be extended (quite painfully, unfortunately) to
$n$-ary classification.  Among kernel methods, they are probably the best
known.  This might be explained by the fact that they enforce sparsity in the
best fit representation. Formally, they ensure that in eq~\ref{eq:bestfit}, a
varying number of $\alpha$'s will be exactly zero.  Sparsity often brings along
a lot of nice properties. For SVMs, most notably, they ensure a better
regularization (intuitively, a fit with less term is less complex than a fit
with many terms, and the number of term is directly related to the number of
non-zero $\alpha$'s), and they induce smaller computational cost. Indeed, the
best fit represented in eq~\ref{eq:bestfit} has as many terms as there are
datapoints in the train dataset. This means that the computational cost at test
time will depend directly on the number of datapoints in the training set. This
might prove computationally costly. SVMs partly solve this problem by ensuring
that some (most) $\alpha$'s are zero, and thus that the best fit will only
depend on a few training datapoints. 
\subsection{Generalities}
Given an input set $\X$, a p.s.d kernel on $\X$, $K$ and a training dataset,
$((x_1, y_1), \dots, (x_n, y_n)) \in (\X \times \pmset)^n$, SVMs solve the
kernel optimization problem eq~\ref{eq:minkernel} with the particular choice of
loss $L(x, y) = \max(0, 1 - x^\top y)$, which is known as the Hinge Loss.

@@

The theory works well, but a use case is always needed. We decided to use the SVM to learn our computer how to recognize handwritten digits. To do so, we create a script in python using the {\it sklearn}\cite{python} library. We used the MNIST database of handwritten digits\cite{mnist} provided by Yann Lecun and well studied. Each of the images (a handwritten digit between 0 and 9) has a given label. The purpose is to create a model using SVM such that the computer can recognize these digits. Some examples of the MNIST dataset are given in Figure \ref{fig:train-set}.

@@

The purpose of this script was to train the computer on 60000 images with their labels. And then, test the model on 10000 images. We could finally compare the predicted labels against the real labels. In order to test the SVM method, we trained the algorithm with different kernel functions:
\begin{itemize}
\item Linear kernel: $K(x_i,x_j) = \langle x_i, x_j \rangle$
\item Polynomial kernel: $K(x_i,x_j) = (\gamma \langle x_i, x_j \rangle + r)^p$, $\gamma$ is a parameter that needs to be found and we decided to keep $r=0$ and we used $p=4$.
\item RBF kernel:  $K(x_i,x_j) = \exp(-\gamma|x_i-x_j|^2)$, $\gamma$ is a parameter that needs to be found.
\end{itemize}

The most important result is the correctness of the prediction. To do this, we take the number of correct prediction and we divide it by the length of the dataset. This gives us a percentage on the correctness of the model. We give in Table \ref{tab:correct} the results for each of the kernels.

\begin{table}[H]
\centering
\begin{tabular}{l|c}
Kernels & Correct prediction \\
\hline
Linear & 94.62\% \\
Polynomial (degree 4) & 97.31\% \\ 
RBF & 98.56 \% \\ 
\end{tabular}
\caption{\label{tab:correct} Percentage of correct prediction for each of the kernels.}
\end{table}

As we can see, the best kernel is the RBF one. It has already been shown in the literature that this kernel is the state-of-the-art in SVM. Another important aspect is to have a look at the predictions more precisely. For this, we use a confusion matrix. This matrix shows the labels (vertical axis) in function of the predictions (horizontal axis). It is interesting to have a look at this matrix to see if there is some bias. (For example, 1 can be often predicted as 7). The confusion matrix for the RBF kernel is given in Table \ref{tab:conf_mat}. As we can see in this table, we don't have any bias. We just show some wrong predictions in Figure \ref{fig:wrong-label}. This figure shows us that some of the images are difficult to label, even for us.

\begin{table}[h]
\centering
\begin{tabular}{cc|c|c|c|c|c|c|c|c|c|c|}
\multicolumn{2}{c}{} & \multicolumn{10}{c}{Predictions} \\
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} & \multicolumn{1}{c}{6} & \multicolumn{1}{c}{7} & \multicolumn{1}{c}{8} & \multicolumn{1}{c}{9}  \\
\cline{3-12}
\multirow{10}{*}{\rotatebox[origin=c]{90}{Labels}}
& 0 & \cca{99}{974} & 0 & 1 & 0 & 0 & 2 & 0 & 1 & 2 & 0 \\ \cline{3-12}
& 1 & 0 & \cca{99}{1128} & 3 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\ \cline{3-12}
& 2 & 4 & 0 & \cca{98}{1017} & 0 & 1 & 0 & 0 & 7 & 3 & 0 \\ \cline{3-12}
& 3 & 0 & 0 & 2 & \cca{99}{997} & 1 & 2 & 0 & 4 & 3 & 1 \\ \cline{3-12}
& 4 & 0 & 0 & 2 & 0 & \cca{99}{968} & 0 & 4 & 0 & 1 & 7 \\ \cline{3-12} 
& 5 & 2 & 0 & 0 & 5 & 1 & \cca{98}{877} & 3 & 1 & 2 & 1 \\ \cline{3-12} 
& 6 & 3 & 2 & 0 & 0 & 2 & 2 & \cca{99}{948} & 0 & 1 & 0 \\ \cline{3-12} 
& 7 & 0 & 3 & 8 & 1 & 1 & 0 & 0 & \cca{98}{1007} & 1 & 7 \\ \cline{3-12} 
& 8 & 3 & 0 & 1 & 3 & 1 & 1 & 0 & 2 & \cca{97}{959} & 4 \\ \cline{3-12} 
& 9 & 2 & 3 & 1 & 6 & 6 & 2 & 1 & 5 & 1 & \cca{100}{982} \\ \cline{3-12} 
\end{tabular}
\caption{\label{tab:conf_mat} Confusion Matrix for the RBF kernel.}
\end{table}

\begin{figure}[H]
\centering
\input{./figures/wrong-label.tex}
\caption{\label{fig:wrong-label} Examples of images that have not been labelled correctly. This number above the pictures are like this: label $\rightarrow$ prediction.}
\end{figure}

\section{Conclusion}

Within this project, we explored a really interesting topic. Indeed, Machine Learning as well as Artificial Intelligence is today's future technology. A lot of start-ups are working on this and applying it to a lot of different fields such as Facial and Emotion Recognition, Language Recognition, Data Mining, etc. Until not so long ago, the SVM were the state-of-the-art. Now, Random Forests and Deep-Learning (Convolutional Neural Networks) are better than this method. But their underlying theory is sometimes not totally understood compared to the SVM. Indeed, Convex Optimization is very known field by many scientist. Developing a model using Convex Optimization to make the computer learn things, such as handwritten digits, is quite amazing.

\begin{thebibliography}{1}

\bibitem{course} \href{http://lear.inrialpes.fr/people/mairal/teaching/2015-2016/MVA/fichiers/mva_slides.pdf}{Slides of the course on "{\it Kernel Methods for Machine Learning}", Inria, Grenoble, France}

\bibitem{paper} Kim, Seung-Jean, et al. "Learning the kernel via convex optimization." {\it Acoustics, Speech and Signal Processing, 2008. ICASSP 2008. IEEE International Conference on.} IEEE, 2008.

\bibitem{svm} \href{https://en.wikipedia.org/wiki/Support_vector_machine}{Wikip\'edia - Support Vector Machine}

\bibitem{python} \href{Library {\it sklearn} for Machine Learning in Python. Functions on SVM/SVC.}{http://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}

\bibitem{mnist} \href{The MNIST Databse of handwritten digits. (Website of Yann Lecun)}{http://yann.lecun.com/exdb/mnist/}

\end{thebibliography}
